---
title: "Anna's Markdown"
author: "Anna Abelman"
date: "3/8/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(RSQLite)
library(tmap)
library(sf)
library(janitor)
library(tidyverse)
#install.packages("USAboundaries")
#install.packages("USAboundariesData", repos = "http://packages.ropensci.org", type = "source")
library(devtools)
#devtools::install_github("ropensci/USAboundariesData")
library(USAboundariesData)
library(USAboundaries)
```



```{r}
my_fire_db <- dbConnect(RSQLite::SQLite(), "FPA_FOD_20170508.sqlite")
```

```{r}
dbListTables(my_fire_db)
```

```{r}
dbListFields(my_fire_db, "Fires")
dbListFields(my_fire_db, "vector_layers")
```

```{r}
fire_data <- dbReadTable(my_fire_db,"fires")
fire_data_geometry <- dbReadTable(my_fire_db,"vector_layers")
```

```{r}
# To query database...
# Eldo_fires <- dbGetQuery(my_fire_db, 'SELECT * FROM fires WHERE "SOURCE_REPORTING_UNIT_NAME" == "Eldorado National Forest"')
# 
# #only look at montana and selecting only a few columns
# mt <- dbGetQuery(my_fire_db, 'SELECT * FROM fires WHERE "STATE" == "MT"') %>% 
#   clean_names() %>% 
#   dplyr::select(fire_name, fire_year, stat_cause_descr, fire_size, latitude, longitude, state, county, source_reporting_unit_name)


# year_2015_tidy <- year_2015 %>% 
#   clean_names() %>% 
#   select(fire_name, fire_year, stat_cause_descr, fire_size, latitude, longitude, state, county, shape)
```


```{r}
#get montana state and counties shapefile 
counties_mt <- us_counties(states = "Montana")
plot(st_geometry(counties_mt))

mt <- read_csv("montana_fire.csv") %>% 
  clean_names() %>% 
  select(-fire_code, -fire_size_class)

mt_sf <- st_as_sf(mt, coords = c("longitude", "latitude"), crs = 4326)

```

```{r}

#considering exporting mt dataframe to make .csv and read in as read_sf --> thoughts?
# ------------------------------------------------------------------------------------------------
#MAIN GOAL IS TO MAKE INTERACTIVE OR AT LEAST USE TMAP
# ------------------------------------------------------------------------------------------------
#every montana fire layered over a map of montana
ggplot()+
  geom_sf(data = counties_mt)+
  geom_point(data = mt, aes (x = longitude, y = latitude),
             alpha = 0.5)

#filter by cause from widget
mt_cause <- mt_sf %>% 
  filter(stat_cause_descr == "Campfire")

# ggplot()+
#   geom_sf(data = counties_mt)+
#   geom_point(data = mt_cause, aes (x = longitude, y = latitude),
#              alpha = 0.5)

#filter by time from widget (have user pick a specific year)
mt_year <- mt %>% 
  filter(fire_year == "2005")

ggplot()+
  geom_sf(data = counties_mt)+
  geom_point(data = mt_year, aes (x = longitude, y = latitude),
             alpha = 0.5)

#filter to search specific fire name in montana
mt_name <- mt %>% 
  filter(fire_name == "GATEWAY")

ggplot()+
  geom_sf(data = counties_mt)+
  geom_point(data = mt_name, aes (x = longitude, y = latitude))

```


```{r}
mt_sf2 <- mt_sf %>% 
  select(fire_size)
#ask about how to have Esri.World Image locked only on MT and have the outline of MT
#ask how to add a gradient of colors for the dots to show difference in fire_area
tmap_mode("view")

map <- tm_basemap("Stamen.TerrainBackground")+
  tm_shape(mt_sf2)+
  tm_dots(label = "fire_size", col = "skyblue",
          size = 0.02)

map

tmap_mode("view")

cause_map <- tm_basemap("Stamen.TerrainBackground")+
  tm_shape(mt_cause)+
  tm_dots(label = "Campfire", col = "orange",
          size = 0.02)

cause_map
  
```

###Next step:

- widget for individuals to pick national forest name and show map of the fires
- search by fire name and show map (or just table of all information aboue the fire)
- causes of the fires (need to condense list since there are a lot of options) --> simple graph showing the frequency of each (ggbeeswarm plot). could also allow the user to pick fire cause and tmap showing dots of that cause in montana
- times series plot showing the containment time over the years --> widget though, pick county/national forest name?


```{r}
#containment time
mt_time <- fire_data %>% 
  clean_names() %>% 
  filter(state == "MT") %>% 
  select(fire_name, fire_year, discovery_date, discovery_time, cont_date, cont_time, source_reporting_unit_name)

#change format of discovery date and continuous date from julian to normal
mt_date <- mt_time %>% 
  mutate(new_dis_date = as.Date(discovery_date, origin = structure(-2440588, class = "Date"))) %>% 
  mutate(new_cont_date = as.Date(cont_date, origin = structure(-2440588, class = "Date")))

#dat$date_time = mdy_hm(paste(dat$date, dat$time))

library(lubridate)

mt_date_combined <- mt_date %>% 
  mutate(final_dis_date = lubridate::ymd_hm(paste(mt_date$new_dis_date, mt_date$discovery_time))) %>% 
  mutate(final_cont_date = lubridate::ymd_hm(paste(mt_date$new_cont_date, mt_date$cont_time))) %>% 
  mutate(interval = difftime(final_cont_date,final_dis_date,units = "min")) %>% 
  mutate(interval = round(interval/60, 2)) %>% 
  drop_na(final_cont_date, final_dis_date)

mt_date_park <- mt_date_combined %>% 
  filter(source_reporting_unit_name == "Kootenai National Forest")

mt_interval_avg <- mt_date_park %>% 
  group_by(fire_year) %>% 
  summarize(
    mean_interval = round(mean(interval),2)
  )

ggplot(data = mt_interval_avg, aes(x = fire_year, y = mean_interval))+
  geom_line(color = "darkgreen") +
  geom_smooth(color = "black",
              size = 0.2,
              linetype = "dashed",
              fill = "lightgreen",
              alpha = 0.2) +
  theme_minimal()


```

```{r}
library(gt)

mt_name %>%  
  gt() %>% 
  tab_header(
    title = "Fire Summary"
  ) %>% 
  cols_label(
    fire_name = "Name of Fire",
    fire_year = "Year",
    stat_cause_descr = "Cause of Fire",
    fire_size = "Fire Area",
    latitude = "Latitude",
    longitude = "Longitude",
    state = "State",
    source_reporting_unit_name = "Name of Reporting Source"
  )
```

```{r}
    mt_sf_test <- mt_sf %>% 
      filter(stat_cause_descr == "Campfire")
  
    leaflet(mt_sf_test)+
      tm_basemap("Stamen.TerrainBackground")+
      tm_shape(mt_sf)+
      tm_dots(label = "Campfire", col = "orange",
              size = 0.02)
```














