---
title: "Shiny App - AA"
author: "Anna Abelman"
date: "3/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(RSQLite)
library(tmap)
library(sf)
library(janitor)
library(tidyverse)
#install.packages("USAboundaries")
#install.packages("USAboundariesData", repos = "http://packages.ropensci.org", type = "source")
library(devtools)
#devtools::install_github("ropensci/USAboundariesData")
library(USAboundariesData)
library(USAboundaries)
```



```{r}
my_fire_db <- dbConnect(RSQLite::SQLite(), "FPA_FOD_20170508.sqlite")
```

```{r}
dbListTables(my_fire_db)
```

```{r}
dbListFields(my_fire_db, "Fires")
dbListFields(my_fire_db, "vector_layers")
```

```{r}
fire_data <- dbReadTable(my_fire_db,"fires")
fire_data_geometry <- dbReadTable(my_fire_db,"vector_layers")
```

```{r}
# To query database...
Eldo_fires <- dbGetQuery(my_fire_db, 'SELECT * FROM fires WHERE "SOURCE_REPORTING_UNIT_NAME" == "Eldorado National Forest"')

#only look at montana and selecting only a few columns
mt <- dbGetQuery(my_fire_db, 'SELECT * FROM fires WHERE "STATE" == "MT"') %>% 
  clean_names() %>% 
  dplyr::select(fire_name, fire_year, stat_cause_descr, fire_size, latitude, longitude, state, county, source_reporting_unit_name)


# year_2015_tidy <- year_2015 %>% 
#   clean_names() %>% 
#   select(fire_name, fire_year, stat_cause_descr, fire_size, latitude, longitude, state, county, shape)
```


```{r}
#get montana state and counties shapefile 
counties_mt <- us_counties(states = "Montana")
plot(st_geometry(counties_mt))

mt_sf <- st_as_sf(mt, coords = c("longitude", "latitude"), crs = 4326)
```

```{r}
#considering exporting mt dataframe to make .csv and read in as read_sf --> thoughts?
# ------------------------------------------------------------------------------------------------
#MAIN GOAL IS TO MAKE INTERACTIVE OR AT LEAST USE TMAP
# ------------------------------------------------------------------------------------------------
#every montana fire layered over a map of montana
ggplot()+
  geom_sf(data = counties_mt)+
  geom_point(data = mt, aes (x = longitude, y = latitude),
             alpha = 0.5)

#filter by cause from widget
mt_cause <- mt_sf %>% 
  filter(stat_cause_descr == "Campfire")

# ggplot()+
#   geom_sf(data = counties_mt)+
#   geom_point(data = mt_cause, aes (x = longitude, y = latitude),
#              alpha = 0.5)

#filter by time from widget (have user pick a specific year)
mt_year <- mt %>% 
  filter(fire_year == "2005")

ggplot()+
  geom_sf(data = counties_mt)+
  geom_point(data = mt_year, aes (x = longitude, y = latitude),
             alpha = 0.5)

#filter to search specific fire name in montana
mt_name <- mt %>% 
  filter(fire_name == "GATEWAY")

ggplot()+
  geom_sf(data = counties_mt)+
  geom_point(data = mt_name, aes (x = longitude, y = latitude))

```


```{r}
mt_sf2 <- mt_sf %>% 
  select(fire_size)
#ask about how to have Esri.World Image locked only on MT and have the outline of MT
#ask how to add a gradient of colors for the dots to show difference in fire_area
tmap_mode("view")

map <- tm_basemap("Stamen.TerrainBackground")+
  tm_shape(mt_sf2)+
  tm_dots(label = "fire_size", col = "skyblue",
          size = 0.02)

map

tmap_mode("view")

cause_map <- tm_basemap("Stamen.TerrainBackground")+
  tm_shape(mt_cause)+
  tm_dots(label = "Campfire", col = "orange",
          size = 0.02)

cause_map
  
```

Next step:

- widget for individuals to pick national forest name and show map of the fires
- search by fire name and show map (or just table of all information aboue the fire)
- causes of the fires (need to condense list since there are a lot of options) --> simple graph showing the frequency of each (ggbeeswarm plot). could also allow the user to pick fire cause and tmap showing dots of that cause in montana
- times series plot showing the containment time over the years --> widget though, pick county/national forest name?

```

